PInSoRo: A Dataset for Deep-learning of Human-Robot Social Interactions

### 2017-06-26 paper.tex
WIP

\todo[inline]{look at other child-child datasets!}
\todo[inline]{mention research on deep-learning to diagnostic autism}

@@

The Plymouth Interacting Social Robots (PInSoRo) Dataset comprises of natural
social interactions arising between children, and recorded in a way suitable
for research methodologies based on machine learning.

As detailed in section~\ref{availability}, this dataset is open and made freely
available to any interested researcher. It is intended to be used to train
machine-learning algorithms (typically, deep recurrent network) to both
recognise and generate complex cognitive and social dynamics occurring in the
wild during social interactions.

@@

The acquisition of such a dataset requires first the definition of a
task to be performed by the participants while they are recorded. The nature of
this task influences in fundamental way the kind of interactions that might be
observed, and thus, learned.  Because they usually target very specific social
or cognitive skills in a tightly controlled manner, traditional socio-cognitive tasks are
either toy scenarios (\ie they do not adequately mirror real-world situations)
or simple, constrained tasks that do not reflect the complexity and dynamics of
real-world interactions. For this dataset, we aim on the contrary at a rich set
of naturally-occurring social interactions, and as such we have designed a novel
task that:

\begin{itemize}
    \item is fundamentally social, \ie the task would make little sense for an
        agent alone;
    \item elicit a large range of interaction situations;
    \item foster rich multi-modal interaction: simultaneous speech, gesture, and gaze
        behaviours are to be observed;
    \item foster non-trivial social dynamics, such as implicit turn-taking.
\end{itemize}

This new experimental task is based on \emph{free play} interactions. Pairs of
children (4-7 years old) are invited to freely draw and interact with items
displayed on an interactive table, telling and acting stories with no explicit
goal set in advance (Fig.~\ref{fig|setup}).

As a result, the interaction space is high-dimensional, along two main axis: the
game-related actions (drawing, manipulating items) and the social interactions
(looking at the other's actions, skink
ocial interaction situation based on
free dramatic play: 
Critically we propose to limit these tasks to
\emph{face-to-face} social interactions, either dyadic or triadic. This constrains the dataset to
a more tractable domain, and should ensure technical feasibility.

, free play social
interactions exhibit a rich set of natural cognitive and social dynamics that
are desirable.

\begin{figure}
    \centering
    \includegraphics[width=0.4\linewidth]{setup-child-child.png}
    \hspace{1em}
    \includegraphics[width=0.4\linewidth]{setup-child-robot.png}
    \caption{The free play social interactions sandbox. Data acquisition is
    performed under two (between subjets) conditions: either children playing together, or one
    child playing with a \emph{non-social} robot.}
    \label{fig|setup}
\end{figure}

To mitigate the technical challenges presented by free play interactions in
their most unconstrained version (like
high dimensionality and uncertainties), we have developped a \emph{free play
sandbox}, Figure~\ref{fig|setup}: a large touchscreen (60cm $\times$ 33cm) is
used as an interactive surface (\emph{sandtray}
paradigm~\cite{baxter2012touchscreen}). Two agents (either a pair of children or
one child and one robot) play together by freely moving on a 2D plane items
(characters and blocks, see Figure~\ref{fig|sandbox}). A background image
depicts a generic empty environment, with different symbolic zones (water,
grass, path, bushes...). The players do not have any particular task to
complete, they are simply invited to freely play.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{sandbox}
    \caption{Example of a game situation: the child and the robot `build'
    together a zoo. Items (animals, characters...) can be dragged over the whole
    play area, while the background picture can be painted over.}
    \label{fig|sandbox}
\end{figure}

>>>>

\todo[inline]{look at other child-child datasets!}
\todo[inline]{mention research on deep-learning to diagnose autism}

@@

The Plymouth Interacting Social Robots (PInSoRo) Dataset comprises of natural
social interactions arising between children, and recorded with machine learning
applications in mind.

As detailed in section~\ref{availability}, this dataset is open and made freely
available to any interested researcher. It is intended to be used to train
machine-learning algorithms (typically, deep recurrent networks) to both
recognise and generate complex cognitive and social dynamics occurring in the
wild during social interactions.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.4\linewidth]{setup-child-child.png}
    \hspace{1em}
    \includegraphics[width=0.4\linewidth]{child-child-env}
    \caption{The free play social interactions sandbox: two children interact in
    Da free-play situation, by drawing and manipulating items on a touchscreen.}
    \label{fig|setup}
\end{figure}

The acquisition of such a dataset requires first the definition of a suitable
task for the participants to perform while they are recorded. The nature of
this task influences in fundamental ways the kind of interactions that might be
observed, and thus, learned. Because they usually target very specific social
or cognitive skills in a tightly controlled manner, traditional socio-cognitive tasks are
usually simple, constrained tasks that do not reflect the complexity and dynamics of
real-world interactions. On the contrary, this dataset aims at capturing a rich set
of naturally-occurring social interactions. As such, a suitable task has to be:

\begin{itemize}
    \item elicit a large range of interaction situations;
    \item foster rich multi-modal interaction: simultaneous speech, gesture, and gaze
        behaviours are to be observed;
    \item fundamentally social, \ie the task would make little sense for an
        agent alone;
    \item foster non-trivial social dynamics, such as implicit turn-taking.
\end{itemize}

We present a new experimental task based on \emph{free play} interactions. Pairs of
children (4-7 years old) are invited to freely draw and interact with items
displayed on an interactive table, telling and acting stories with no explicit
goal set in advance (Fig.~\ref{fig|setup}).

In order to constraint the dataset to a tractable domain, the task is limited to
a dyadic \emph{face-to-face} interaction.  The resulting interaction space is
nevertheless high-dimensional, with two main axis: the game-related actions
(drawing, manipulating items, telling stories) and the social interactions
(monitoring the other's actions, discussing and agreeing on joint actions,
helping behaviours, etc.).

The \emph{freeplay sandbox} follows the sandtray
paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
multitouch) is used as an interactive surface (\emph{sandtray}). The two children play together
by freely moving interactive items (Fig.~\ref{fig|sandbox}) on the surface. A background image
depicts a generic empty environment, with different symbolic zones (water,
grass, path, bushes...). The players do not have any particular task to
complete, they are simply invited to freely play.
is implemented wit To mitigate the technical challenges presented by free play interactions in
their most unconstrained version (like
high dimensionality and uncertainties), we have developped a \emph{free play
sandbox}, Fig.~\ref{fig|setup}:

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{sandbox}
    \caption{Example of a game situation: the two children `build'
    together a zoo. Items (animals, characters...) can be dragged over the whole
    play area, while the background picture can be painted over by picking a
    colour.}
    \label{fig|sandbox}
\end{figure}

@@
